### 2014-09-19 paper/include/techniques.tex
Wrote a little

\section{Intelligent Refresh Techniques} 
\label{sec:int}

Say a little about every technique.
\subsection{Access Recency}

\subsection{Retention}

\subsection{Tolerance}

\subsection{Validity}