### 2014-09-19 paper/include/accessrecency.tex

\section{Access Recency}
\label{sec:acc}

Papers that use this technique:

smart
refrint
P. Emma et al.