### 2014-09-19 paper/include/accessrecency.tex
cite added

\section{Access Recency}
\label{sec:acc}

Papers that use this technique:

smart\cite{smart}

refrint\cite{refrint}

P. Emma et al.