### 2014-09-19 paper/include/retention.tex

\section{Retention}
\label{sec:ret}
Papers that use this technique:

DTail\cite{dtail}

VRA

RAIDR\cite{raidr}

RIO\cite{rioparis}