### 2014-09-19 paper/include/tolerance.tex

\section{Tolerance} 
\label{sec:tol}

Papers that use this technique:

Flikker\cite{flikker}