### 2014-09-19 paper/include/accessrecency.tex
explained techniques

\section{Access Recency}
\label{sec:acc}

Papers that use this technique:

smart\cite{smartrefresh}

By employing a time-out counter for each memory row of a DRAM module, all the unnecessary periodic refresh operations can be eliminated. The basic concept behind our scheme is that a DRAM row that was recently read or written to by the processor (or other devices that share the same DRAM) does not need to be refreshed again by the periodic refresh operation, thereby eliminating excessive refreshes and the energy dissipated.

refrint\cite{refrint}

P. Emma et al.