### 2014-09-19 paper/include/retention.tex

\section{Retention}
\label{sec:ret}
Papers that use this technique:

DTail\cite{dtail}

VRA

RAIDR\cite{raidr}, from 2006...
Software approach that can exploit off-the-shelf DRAMs to reduce refresh power to vanishingly
small levels approaching non-volatile memory. The key idea is to favor longer-retention pages over shorter-retention pages when allocating DRAM pages. This allows selecting a single refresh period that depends on the shortest-retention page among populated pages,
instead of the shortest-reten tion page overall.

RIO\cite{rioparis}

SECRET\cite{secret}
Previous works adopt conventional ECC (Error Correcting Code) to correct retention errors. These approaches introduce significant area and energy overheads. In this paper, we propose a novel error correction framework for retention errors in DRAMs, called SECRET (Selective Error Correction for Refresh Energy reducTion).

Energy Optimization in 3D MPSoCs with Wide-I/O DRAM Using Temperature Variation Aware Bank-wise Refresh

We improve DRAM refresh power by considering the lateral and vertical temperature
variations in the 3D structure and adapting the per-DRAM-bank refresh period accordingly.