### 2014-09-21 paper/include/retention.tex

\section{Retention}
\label{sec:ret}

Zehan Cui et al. \cite{dtail}

RAIDR  by Jinghui Liu et al. \cite{raidr}

RIO is a technique proposed by Seungjae Baek et al. \cite{rioparis}

With SECRET Chung-Hsiang Lin et al. \cite{secret} masks bad cells using Error Correction Pointers (ECP). The bad cells are maped with a one time profiling done before first system use and all cells that dont meet the retention time requirements gets a ECP. A ECP is basiclly a pointer to another cell that is healthy. The framwork which is implemented inside the memory controller checks wheter there is any valid ECP for every address accessed. If there is, it replaces the bad cell(s) data with the data from ECP cell instead.

This method gives a high energy saving of up to 87.2\% in refresh power and a reduction of 18.57\% in total DRAM power consumption.

RAIDR
Software approach that can exploit off-the-shelf DRAMs to reduce refresh power to vanishingly
small levels approaching non-volatile memory. The key idea is to favor longer-retention pages over shorter-retention pages when allocating DRAM pages. This allows selecting a single refresh period that depends on the shortest-retention page among populated pages,
instead of the shortest-reten tion page overall.