### 2014-09-30 paper/include/smart_refresh.tex
cleaned

The basic concept behind our scheme is that a memory line that has been recently read out or written to does not need to be refreshed again by the periodic refresh mechanism.

Uses a time-out counter for each memory line. The counter is set on a read/write/refresh and dicards refreshes until the timer hits zero. The counter is associated to a bank/line pair. The counters are stored in the memory controller. One counter consists of 2 SRAM bits. More bits give higher granularity and thus also better performance because the refresh is made less often.

This scheme will be most effective if all lines are accessed before a refresh is needed. If no accesses are made = the entire working set fits into cache, Smart Refresh becomes CBR (CAS before RAS). Only refresh one line at a time = inefficient, so they have made it possible to turn the technique off.

Implemented so only one line per each logical segments are deremented at the same time, which leads to that the refreshes are evenly distributed over time. (I.e. no burst refresh.)

Overhead: 48KB for 2GB and 768 KB for 32GB = 0,0024\% overhead.

Average reduction of refreshes by 59.3\% for 2GB and 52.3\% energy saving regarding the DRAM refresh. No performance degradation.

Employed its solution to 2 GB ram. What about todays sizes? How effective is the technique now? - Only works effective if large parts the memory is acccessed, i.e. large part of the memory has to be valid to begin with. Is it so today?