### 2014-10-02 paper/include/cream.tex
made some placeholders for new subsections

%CREAM
Do not know where to place it. (RTAV)

\todo[inline]{Ccopy-paste:}
To mitigate the refresh penalty, a Concurrent-REfresh-Aware Memory sys-tem (CREAM) is proposed in this work so that memory access and refresh can be served in parallel. The proposed CREAM architecture distinguishes itself with the following key contributions: (1) Under a given DRAM power budget,
sub-rank-level refresh (SRLR) is developed to reduce refresh power and the saved power is used to enable concurrent memory access; (2) sub-array-level refresh (SALR) is also devised to effectively lower the probability of the conflict between memory access and refresh; (3) In addition, novel sub-array level refresh scheduling schemes, such as sub-array round-robin and dynamic scheduling, are designed to further improve the performance