### 2014-10-08 paper/include/flikker.tex
wrote, added TODOs and sources

%flikker
Tolerance!\cite{flikker}
\todo[inline]{Add description.}