### 2014-10-09 paper/include/raidr.tex

RAIDR by J. Liu et al. \cite{raidr} divides all memory lines into different groups, called bins, based on the cell with the shortest retention time in the line. In the basic configuration there are two bins, one with a refresh rate of \(64ms - 128ms\) and one with \(128ms - 256ms\). Lines in the first bin gets refreshed every \(64ms\), lines in the second bin every \(128ms\), and lines not in any bin gets refreshed at the default refresh rate which is set to \(256ms\). The implementation is done inside the memory controller and to minimize the storage needed for keeping track of bin members they use Bloom filters. Thanks to this only 1.25 KB storage is needed for a 32 GB DRAM with the default configuration. The storage requirement can however vary depending the amount of bins used. RAIDR also accommodate for temperature variations by scaling all bins refresh rates using a timer. With RAIDR a performance improvement of 4.1\% and a decrease in power consumption by 8.3\% is achieved in a system with 32GB DRAM compared to auto-refresh.