### 2014-10-10 lab/report/main.tex
modified the make script

\documentclass[a4paper]{report}

\include{packages}

%MACROS---
\newcommand{\reffig}[1]{Fig. \ref{#1}}

\begin{document}
\section{test}
heeeej!!! :)
\end{document}