### 2014-10-11 lab/report/main.tex
hrj

\documentclass[a4paper]{article}

\include{packages}

%MACROS---
\newcommand{\reffig}[1]{Fig. \ref{#1}}

\title{Report on Lab 1 \& Lab 2 in EDA282, Parallel Computer Organization and Design}
\date{October 2014}
\author{Dan Larsson, Jonas Hemlin
\\Master Program of Embedded Electronic Systems Design\\
Chalmers University of Technology, Gothenburg, Sweden\\
Email: \{larsdan, jhemlin\}@student.chalmers.se}

\begin{document}
\maketitle
\input{include/lab1}
\input{include/lab2}
\input{include/discussion}
\end{document}