### 2014-10-11 lab/report/include/introduction.tex

\section{Introduction}
\label{sec:int}
These laboratory assignments aim to increase our knowledge of different cache coherence protocols. The goal is to simulate cache coherence protocols on a modeled cache coherent multi-core system using diverse benchmarks. Three cache coherence protocols are targeted; MSI, MESI, and MESI-MG, where the last protocol is an modified version of MESI optimized for migratory sharing. The system model have a shared Last Level Cache (LLC), and each processor have a private L1 cache. Snooping based write-invalidate cache coherence protocols are used by the private L1 caches to ensure coherence, and a simple bus stand for the communication between LLC and private caches. The system is modeled by the Pin tool MultiCacheSim and Pin.