### 2014-10-12 lab/report/main.tex
changed margins

\documentclass[a4paper]{article}

\usepackage[margin=2cm]{geometry}
\include{packages}
\usepackage{color}
%MACROS---
\newcommand{\reffig}[1]{Fig. \ref{#1}}
\newcommand\xput[2][0.5]{%
    \rule{#1\linewidth}{0pt}\makebox[0pt][c]{#2}\hfill}

\title{Report on Lab 1 \& Lab 2 in EDA282, Parallel Computer Organization and Design}
\date{October 2014}
\author{Dan Larsson, Jonas Hemlin
\\Master Program of Embedded Electronic Systems Design\\
Chalmers University of Technology, Gothenburg, Sweden\\
Email: \{larsdan, jhemlin\}@student.chalmers.se}

\definecolor{mygreen}{rgb}{0,0.6,0}
\definecolor{mygray}{rgb}{0.5,0.5,0.5}
\definecolor{mymauve}{rgb}{0.58,0,0.82}

\lstset{ 
	language=C,
	backgroundcolor=\color{white},
	basicstyle=\scriptsize,        % the size of the fonts that are used for the code
	breakatwhitespace=false,         % sets if automatic breaks should only happen at whitespace
	breaklines=true,                 % sets automatic line breaking
	captionpos=t,                    % sets the caption-position to bottom
	commentstyle=\color{mygreen},    % comment style
	deletekeywords={},
	morekeywords={pragma,schedule,reduction,omp,single,private,shared, parallel},            
	extendedchars=true,              % lets you use non-ASCII characters; for 8-bits encodings only, does not work with UTF-8
	frame=single,                    % adds a frame around the code
	keepspaces=true,                 % keeps spaces in text, useful for keeping indentation of code (possibly needs columns=flexible)
	keywordstyle=\color{blue},       % keyword style
	numbers=left,                    % where to put the line-numbers; possible values are (none, left, right)
	numbersep=5pt,                   % how far the line-numbers are from the code
	numberstyle=\tiny\color{mygray}, % the style that is used for the line-numbers
	rulecolor=\color{black},         % if not set, the frame-color may be changed on line-breaks within not-black text (e.g. comments (green here))
	showspaces=false,                % show spaces everywhere adding particular underscores; it overrides 'showstringspaces'
	showstringspaces=false,          % underline spaces within strings only
	showtabs=false,                  % show tabs within strings adding particular underscores
	stepnumber=2,                    % the step between two line-numbers. If it's 1, each line will be numbered
	stringstyle=\color{mymauve},     % string literal style
	tabsize=2                       % sets default tabsize to 2 spaces                   % show the filename of files included with \lstinputlisting; also try caption instead of title
}

\begin{document}
\maketitle

\input{include/introduction}
\input{include/lab1}
\input{include/lab2}
\input{include/discussion}
\end{document}