### 2014-10-17 paper/include/flikker.tex

%flikker
Tolerance!\cite{flikker}
\todo[inline]{Drafty below.}

Application-level technique.
Targets smartphones on the first hand.

Modifies the MC.
Extends Partial-Array Self Refresh Mode (PASR), which was introduced 2002 in mobile DRAM.

Modifies SW: Applications, the runtime system and the OS.