### 2014-10-23 paper/include/discussion.tex

\section{Discussion} 
\label{sec:disc}

\todo[inline]{Drafty text below!}

\textbf{Mention some smart scheduling techniques}
herpa herpa

CREAM CREAM \cite{cream}

Another proposed technique is ESKIMO by Isen and John \cite{eskimo}, which focus at reducing total DRAM power through optimizing the accesses to the DRAM. By using one bit of a cache row to denote whether it hold nonsense data, it becomes possible to resolve some memory accesses in the caches, which otherwise would have needed to access DRAM. A memory region is regarded as nonsense if it has been deallocated, or allocated but not written to. The ISA has to be extended to provide the allocation and deallocation information. If a block has to be replaced in a write-through cache and the replaced block is dirty but also deallocated, the cache can ignore writing the replaced block to memory. Similarly, a write miss to a newly allocated area results in a read access that can also be ignored. Combining ESKIMO with the selective refresh implementation proposed by Ohsawa \cite{ohsawa}, $39\%$ of total DRAM power was saved on average.

\textbf{Broadly compare the techniques}
\begin{table*}[t]
	\caption{\label{tbl:summary}Summarized stuff.}
        \begin{center}
			\normalsize
			\begin{tabular}{ l r r r r r r}
				\textbf{Technique} & \parbox{2cm}{ \textbf{Refresh \linebreak information*}} & \textbf{Modifies} & \parbox{1.7cm}{ \textbf{Refresh \linebreak reduction}} & \parbox{2.2cm}{ \textbf{DRAM power \linebreak reduction}} & \parbox{1.2cm}{\textbf{Storage \newline Overhead}} & \parbox{1.5cm}{\textbf{Performance impact}} \vspace{0.05cm} \\
				\hline
				\textit{Smart Refresh} & A & MC & 59\% & $12.13\%$ & $0.0048\%$  & Unknown  \\
				\textit{Refrint} & A & MC, (?) & N/A & N/A  & $0.005\%$  & N/A  \\
				\hline
				\textit{RAIDR} & R & MC & $74.6\%$ & $8.3\%$  & $0.031\%$  & $4.1\%$  \\
				\textit{DTail} & V, R & MC, DDRx & $88.2\% \to 98.9\% $ & $< 23\%$  & $0.006\%$ & R:  \\
				\textit{RIO} & R & OS & $87.5\%$ & Unknown  & $0.1\%$  & $4.5\%$ \\
				\textit{SECRET} & R & MC & $87.5\%$ & $18.6\%$  & $>0.01\%$  & $\pm 1.4\%$  \\
				\hline
				\textit{PARIS} & V & MC & Unknown & $-41.7\%$ DTail-V  & $0.0015\%$  & $-4.4\%$ DTail-V  \\
				\hline
				\textit{Flikker} & T & MC, OS, Apps & Unknown & $20\% \to 25\%$  & $0.005\%$  & $-1\%$   \\
				\textit{Sparkk} & V, T  & MC, OS, Apps &  $+50\%$ Flikker & Unknown & Unknown  & Unknown \\
				\hline
			\end{tabular}
		\end{center}
\end{table*}

Smart refresh
Result shows that the refresh reduction was in average 59\%
and refresh power consumption savings was on average 52\%,
all without any performance degradation. 12.13\% reduction of total DRAM
The memory
footprint of the modifications is 0.0048\% of total device
capacity when three-bits counters are used assuming a DRAM
row size of 8 KB.

Refrint
N/A

Flikker
The partitioning scheme can potentially lower performance as
locality of data elements changes, but the found performance
degradation was 1\% or less. No loss of reliability was ex-
perienced, and the technique saved between 20-25\% of total
DRAM power consumption depending on application.
In simulations, N was two bits and Count
was 5 bits, giving a total overhead of 0.005\%, with a row size
of 8 KB.

Sparkk
Sparkk produces higher peak SNR at half the
average refresh rate used for Flikker.

Modifies applications:
Flikker, Sparkk

Modifies OS/runtime system:
ESKIMO, Flikker, Sparkk.

Modifies MC:
All but RIO and PARIS.

Only in SW:
RIO, PARIS.

ROR doesn't work with $<$ 32 GB, cant keep up the req. speed.

\textbf{Advantages and disadvantages for each approach}
-What is modified? / Where is stuff stored?
-How much knowledge does the techniques have about the running software?
-Is the approach's benefits application-depended?
-PARIS could have used CBR if sREF was implemented (as Dtail does), but this extends DDRx instruction set...

Smart Refresh:
Employed its solution to 2 GB ram. What about todays sizes? How effective is the technique now? - Only works effective if large parts the memory is accessed, i.e. large part of the memory has to be valid to begin with. Is it so today?

Smart Refresh \& Refrint:
Refrint is somewhat a extension of Smart Refresh. Smart Refresh's features can be laid somewhat equal to Refrint's time-based polices, whereas the data-based polices have no equivalence in Smart Refresh.

They both modify the memory controller, and almost in the same way. But Refrint has no timers that is getting decremented the whole time, and instead copies a global `timer' to the line associated bits upon access. Of course Refrint's data-based polices also adds logic to the memory controller.

Discuss energy savings. Hard to do eDRAM vs DRAM...

\textbf{Which techniques can be combined?}

... any of these techniques can be combined with `scheduling real accesses together with refresh ops' (can them???)

\textbf{3D stacked DRAM discussion}
-Will any approach be benefited / dis-benefited by it?

-Which technique seems future proof?