### 2014-10-23 paper/include/summary.tex
Changed name on conclussion

\section{Summary} 
\label{sec:sum}
SumSumSUm!! my precious!