### 2014-10-23 paper/include/summary.tex
summary! woop woop

\section{Summary} 
\label{sec:sum}
We have surveyed nine different techniques for lowering the number of refresh operations performed in a modern DRAM system. Because the portion of power consumption coming from refresh operations in DRAM is high and increasing, these techniques are both topical and highly considerable. To give a context to the techniques surveyed the basic operation of DRAM is elaborated on.

All the techniques surveyed fits into one of the approaches based on four types of information, Retention Time, Access Recency, Data Tolerance, and Data Validity, mentioned in the paper. The relative performance of the techniques is compared, and some projections for the future concerning higher density and bigger capacity DRAMs are discussed.