### 2014-10-23 paper/include/discussion.tex
disc text

\section{Discussion} 
\label{sec:disc}

% \begin{table*}[t]
% 	\caption{\label{tbl:summary}Summarized stuff.}
%         \begin{center}
% 			\normalsize
% 			\begin{tabular}{ l r r r r r r}
% 				\textbf{Technique} & \parbox{2cm}{ \textbf{Refresh \linebreak information}} & \textbf{Modifies} & \parbox{1.7cm}{ \textbf{Refresh \linebreak reduction}} & \parbox{2.2cm}{ \textbf{DRAM power \linebreak reduction}} & \parbox{1.2cm}{\textbf{Storage \newline Overhead}} & \parbox{1.5cm}{\textbf{Performance impact}} \vspace{0.05cm} \\
% 				\hline
% 				\textit{Smart Refresh} & A & MC & 59\% & $12.13\%$ & $0.0048\%$  & Unknown  \\
% 				\textit{Refrint} & A & MC, (?) & N/A & N/A  & $0.005\%$  & N/A  \\
% 				\hline
% 				\textit{RAIDR} & R & MC & $74.6\%$ & $8.3\%$  & $0.031\%$  & $4.1\%$  \\
% 				\textit{DTail-R} & R & MC, DDRx & $87.9\%$ & $\approx 23\%$  & $0.0045\%$ & $\approx$ \textit{RAIDR} \\
% 				\textit{RIO} & R & OS & $87.5\%$ & Unknown  & $0.1\%$  & $4.5\%$ \\
% 				\textit{SECRET} & R & MC & $87.5\%$ & $18.6\%$  & $\approx 0.01\%$  & $\pm 1.4\%$  \\
% 				\hline
% 				\textit{DTail-V} & V & MC, DDRx, (OS) & $\approx 10\% \to 90\% $ & $41.7\%$ \textit{PARIS}  & $0.0015\%$ & $4.4\%$ \textit{PARIS}  \\
% 				\textit{PARIS} & V & MC, OS & $\approx 10\% \to 90\%$ & Unknown  & $0.0015\%$  & Unknown \\
% 				\hline
% 				\textit{Flikker} & T & MC, OS, Apps & Unknown & $20\% \to 25\%$  & $0.005\%$  & $-1\%$   \\
% 				\textit{Sparkk} & V, T  & MC, OS, Apps &  $50\%$ \textit{Flikker} & Unknown & Unknown  & Unknown \\
% 				\hline
% 			\end{tabular}
% 		\end{center}
% \end{table*}
In the following section we compare all surveyed techniques more thoroughly and present some of our thoughts on their future potential.

\subsection{Advantages and disadvantages of each approach}
First of, we believe that the footprint of each surveyed technique is negligible when DRAM capacity increases. This as no technique has a larger storage overhead than $0.1\%$ of DRAM capacity.

%\subsubsection*{Access Recency} 
Access Recency techniques becomes less effective when DRAM capacity scales, as it is probable that the access rate per row will decrease. Among the approaches, this is one of the easier to implement, as it essentially only needs to track DRAM accesses and count the time between them. Techniques based on the other approaches can achieve the same or better results even when the access rate is low.

%\subsubsection*{Retention Time}
Using R as the refresh data gives a high efficiency independent of DRAM capacity, running applications, and memory utilization. A downside is that retention time profiling of the DRAM has to be performed in conjunction with first system startup. As retention time is highly dependent on temperature variations and has to be accounted for, which some techniques do; RAIDR, and RIO.

%\subsubsection*{Validity} 
The V approach scales well, as it is more probably that more rows becomes invalid when DRAM capacity increases. It also produce high efficiency in many common systems. However, support on a OS level is needed to gather the V data, which increases the complexity.

%\subsubsection*{Data Tolerance} 
The efficiency of T based techniques is dependent of memory utilization besides of data with low criticality, as T do not target invalid memory areas. Using T information can yield a high decrease of refresh operations, but heavy modifications in several layers are required.

\subsection{Modifications required and future potential}

%\todo[inline]{describe the table}
%-no area info

The method used for refresh can affect the possibility of a technique to scale. Techniques that relies on ROR can have problems with large DRAM capacities. This as the number of rows will be to large for ROR to refresh within the maximum refresh period. One example of this is PARIS, which fails to work on a system with \textit{32~GB} of memory \cite{dtail}.

Techniques which only require OS modifications is easier to adopt than those who need other changes. RIO is a good example of a technique which can be adopted after light kernel updates. The potential of RIO is significant compared to the modifications needed, as it reduces refreshes to the same degree as R based techniques that need hardware modifications at the cost of higher storage overhead. Moreover, all other surveyed techniques require specific support from at least the MC.

Among the techniques which modifies the MC, there are one group that in addition to their logic also have large tables in the MC; Smart Refresh, Refrint, RAIDR, and PARIS. Another group stores it's data in the DRAM instead of the MC; SECRET, and DTail. The advantage of keeping the refresh data in DRAM is lower MC area overhead and cheaper memory, whereas the downside is that the data has to be cached or prefetched to the MC to mitigate the longer access time of DRAM. DTail has additional need of a extension to the DDRx standard, which can complicate the adoption of the technique.

The techniques that requires the heavies modifications are those based on T. Even so, their potential is good and Sparkk's total DRAM power reduction is high, but the changes needed will not be transparent for the application programmer, who has to learn the concept of approximate data in order for the technique to be efficient. This is maybe to much to ask, therefor we do believe that T based techniques has lower future potential than R and V based techniques.

We believe that the R and V approaches is the most promising for the future. This as techniques based on A, even though they are quite simple to implement, the trend is to increase the DRAM capacity and cache structures, which speaks against the usage of this approach. But of course, in systems that does not follow these trends, Access Recency can have a chance to be adopted. Such systems could for example be small embedded systems or large shared memory systems with small or no cache and many cores.

3D stacked DRAM will affect all approaches. The higher temperature increases the refresh rate and stretches the demands for techniques using ROR to lower the refresh rate in order to keep up. The temperature will also be more uneven across the device due to the stacking, which give higher requirements for tracking the variations and adapt to them. One technique not surveyed in this paper, proposed by Sadri et al. \cite{tempaware}, tracks the temperature in different regions of a 3D DRAM and adapts the refresh rate accordingly in these areas.

\subsection{Combining techniques}

To achieve a larger DRAM power reduction, techniques from different approaches can be combined\footnote{Or you just shut off the computer and enjoy the autumn rain}. An example of this is RIO+PARIS, which increases the refresh reduction to $93.8\%$. If DTail-RV is used instead, the refresh reduction increases to $98.9\%$. DTail could be extended with an arbitrary T data acquisition technique for further improvements. All surveyed techniques target to reduce refresh power, but to achieve the greater goal of minimizing total DRAM power, smart scheduling of all DRAM accesses could advantageously be incorporated.

For example has Isen and John proposed the smart scheduling technique ESKIMO \cite{eskimo}, which focus at reducing total DRAM power through optimizing the accesses to the DRAM. By using one bit of a cache row to denote whether it hold nonsense data, it becomes possible to resolve some memory accesses in the caches, which otherwise would have needed to access DRAM. A memory region is regarded as nonsense if it has been deallocated, or allocated but not written to. The ISA has to be extended to provide the allocation and deallocation information. If a block has to be replaced in a write-through cache and the replaced block is dirty but also deallocated, the cache can ignore writing the replaced block to memory. Similarly, a write miss to a newly allocated area results in a read access that can also be ignored. Combining ESKIMO with the selective refresh implementation proposed by Ohsawa \cite{ohsawa}, $39\%$ of total DRAM power was saved on average.