### 2014-10-31 paper/include/summary.tex
some summary. DTAIL POWER!!!!!

\section{Summary} 
\label{sec:sum}
*conducted survey

*dram power importance

*approaches

*our ideas

Of the techniques surveyed DTail by Cui et al. stands out by showing promise for being a platform on which combining many kinds approaches is possible. Even as some of the approaches brought up in this paper shows great potential for tackling the problem with refresh power, it is important to note that they are not the only possible way. Other ways not surveyed, like scheduling of DRAM access, could also play an important role.