### 2014-10-31 paper/include/summary.tex
hej

\section{Summary} 
\label{sec:sum}
We have conducted a survey comprising of several proposed techniques to optimize DRAM refresh power. The survey has been performed  These techniques are based on one or several of the four approaches; Access Recency, Retention Time, Data Tolerance and Validity. DRAM refresh power is an increasing part of computing systems and to be able to mitigate this trend