### 2014-10-31 paper/include/summary.tex

\section{Summary} 
\label{sec:sum}
Due to the increasing importance of lowering power consumption in modern computer systems and as DRAM refresh power is an increasing part of computing systems many techniques targeting DRAM power has been proposed. We have conducted a survey comprising of several of these proposed techniques that optimize DRAM refresh power. The surveyed techniques are based on one or several of the four approaches, Access Recency, Retention Time, Data Tolerance and Validity. The approaches which we believe have higher future potential are Retention Time and Validity. This opinion is based on the fact that those approaches are more capable to handle the DRAM trends and provide good power reduction potential compared to the approaches' design complexity.

Findings from the survey shows that the best way to optimize DRAM refresh is to combine several of the approaches. Of the techniques surveyed, DTail by Cui et al. stands out by showing promise by providing a framework where it is possible to combine different approaches. Even though some of the approaches brought up in this paper shows great potential for tackling the problem with refresh power, it is important to note that their use only are a part of the solution. Other ways not surveyed, like scheduling of DRAM access, will also play an important role.