### 2014-10-31 paper/include/reducing_refresh_power.tex

\section{Reducing refresh power} 
\label{sec:red}
There are several approaches to decrease the refresh power by reducing unnecessary refreshes and these can be categorized depending on which sort of information the scheduling decisions are based upon. The four types of information are: access recency, cell retention time, error tolerance of the data, and line validity. This categorization follows the taxonomy proposed by Cui et al. \cite{dtail}. These four approaches are explained further in \ref{sec:red:app}, whereas techniques based on the approaches are summarized in \ref{sec:red:tech}.

\subsection{Approaches to reduce refresh power}
\label{sec:red:app}

\subsubsection*{\textbf{Access Recency (A)}}
As mentioned, a DRAM access is performed in the same way as a refresh, which makes it possible to delay the next refresh operation to an accessed row without endangering the stored data. All accessed rows within the maximum refresh period result in that a refresh can be postponed. This implies that if the system seldom accesses the DRAM data, or only a small portion of it, the approach's gain diminishes.

\subsubsection*{\textbf{Retention Time (R)}}
The need for refresh is as mentioned caused by leakage in the memory cells. Retention time is a measurement of the time it takes before a cell loose its data if not refreshed (recharged). Variations in the manufacturing process causes a exponentially distributed fluctuation \cite{katayama} in the retention time of the DRAM cells. To accommodate for the cells with the lowest retention time, manufacturers produce chips with average retention times much higher than possible minimum. This gives room for optimization using variable refresh rate for different cells. The retention time is also highly sensitive to temperature variations and becomes drastically lower at higher temperatures. For techniques using this approach the biggest advantage is that the improvements are independent of workload, except for a potential rise in temperature which would lower the retention time.

\subsubsection*{\textbf{Data Tolerance (T)}}
In the other approaches all data loss is seen as intolerable. In this approach data is categorized depending on how critical it is, which yields that cells containing data with a lower criticality may be refreshed more seldom. Applications such as approximate computations, games, or media processing can tolerate some bit errors while still producing a acceptable result. Using this approach, the energy and performance savings achieved is highly depending on what kind of applications most commonly are run on the system.

\subsubsection*{\textbf{Validity (V)}}
In a system that does not utilize all the available memory space it is unnecessary to refresh parts not used. Therefore, cells that do not keep valid data do not need to be refreshed. The allocation and deallocation of the cells have to be tracked to get the information needed for implementation. This approach is highly sensitive to memory utilization of the system and its effects on energy consumption and performance improvements will be diminishing on systems with high workload.

\subsection{Techniques to reducing refresh power}
\label{sec:red:tech}

\subsubsection*{\textbf{Smart Refresh}}
\label{par:smartrefresh}
\input{include/smart_refresh}

\subsubsection*{\textbf{Refrint}}
\label{par:refrint}
\input{include/refrint}

\subsubsection*{\textbf{RAIDR}}
\label{par:raidr}
\input{include/raidr}

\subsubsection*{\textbf{DTail}}
\label{par:dtail}
\input{include/dtail}

\subsubsection*{\textbf{RIO and PARIS}}
\label{par:rioparis}
\input{include/rioparis}

\subsubsection*{\textbf{SECRET}}
\label{par:secret}
\input{include/secret}

\subsubsection*{\textbf{Flikker}}
\label{par:flikker}
\input{include/flikker}

\subsubsection*{\textbf{Sparkk}}
\label{par:sparkk}
\input{include/sparkk}